Quantum Mechanics Assignment Two

### 2016-01-20 quantum/latex/a2.tex
bla

\begin{align*}
  P_{forb} &= \int_{\infty}^{-x_0} \psi_0^*\psi_0 dx + \int_{x_0}^{\infty} \psi_0^*\psi_0 dx\\
  &= \int_{\infty}^{-x_0} \sqrt{\frac{m\omega}{\pi\hbar}}e^{-m\omega x^2 / \hbar}dx
  + \int_{x_0}^{\infty} \sqrt{\frac{m\omega}{\pi\hbar}}e^{-m\omega x^2 / \hbar} dx\\
  &= \sqrt{\frac{m\omega}{\pi\hbar}} \left(\frac{-\hbar}{m\omega}\right)\left(
    \frac{1}{x}e^{-m\omega x^2 / \hbar} \Bigg|_{-\infty}^{-\sqrt{\frac{\hbar}{m\omega}}} +
    \frac{1}{x}e^{-m\omega x^2 / \hbar} \Bigg|_{\sqrt{\frac{\hbar}{m\omega}}}^{\infty}\right)\\
  &= \frac{2}{\sqrt{\pi}} = 0.41\\
\end{align*}

@@

%% \begin{align*}
%%   c_0 &= \int_{-\infty}^{\infty} \phi_0^*\psi dx\\
%%   &= A\int_{-\infty}^{\infty} \left(\frac{m\omega}{\pi\hbar}\right)^{1/2}e^{-m\omega x^2/\hbar}
%%   \left(1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
%%   2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/(2\hbar)}dx\\
%%   &= 2\\
%%   c_1 &= A\int_{-\infty}^{\infty} \frac{2\sqrt{\frac{m\omega}{\hbar}x}}{\sqrt{2}}
%%     \left(\frac{m\omega}{\pi\hbar}\right)^{1/2} \left(1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
%%     2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/\hbar}dx\\
%%  &= -\frac{3}{\sqrt{2}}\\
%%  c_2 &= A\int_{-\infty}^{\infty} \frac{\frac{4m\omega}{\hbar}x^2 - 2}{\sqrt{8}}
%%     \left(\frac{m\omega}{\pi\hbar}\right)^{1/2} \left(1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
%%     2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/\hbar}dx\\
%%  &= \sqrt{2}\\
%% \end{align*}

%% From this we can see that $\psi(x,0) = \sqrt{\frac{21}{2}}
%% \left(2\ket{0} + -\frac{3}{\sqrt{2}}\ket{1} + \sqrt{2}\ket{2}\right)$. It is then easy to compute
%% the expectation value: \\

%% \begin{align*}
%%   <E> &= \sigma \hbar\omega\left(n+\frac{1}{2}\right)|c_n|^2\\
%%   &= \frac{2 + \frac{27}{4} + 5}{\frac{21}{2}} = 1.3 eV\\
%% \end{align*}
\section{5: 2D QHO}
To take the QHO to two dimensions, we use the following Hamiltonian:

@@

We notice that none of the terms in this Hamiltonian are coupled, so we break it
up into two Hamiltonians:

\begin{align*}
  H_x &= \frac{\hat{p}_x^2}{2m} \frac{1}{2}m\omega^2x^2)\\
  H_y &= \frac{\hat{p}_y^2}{2m} \frac{1}{2}m\omega^2y^2)\\
  H &= H_x + H_y\\
\end{align*}

our energy eigenvalue equation then takes the form:

\begin{align*}
  \left(\hat{H}_x + \hat{H}_y\right)X(x)Y(y) &= (E_x + E_y)X(x)Y(y)\\
\end{align*}

@@

\begin{align*}
  \psi_0(x,y) &= \left(\frac{m\omega}{\pi\hbar}\right)
  e^{-m\omega\left(x^2+y^2\right)/2\hbar}\\
\end{align*}

>>>>

\begin{align*}
  P_{forb} &= \int_{\infty}^{-x_0} \psi_0^*\psi_0 dx + \int_{x_0}^{\infty} \psi_0^*\psi_0 dx\\
  &= 2\int_{x_0}^{\infty} \sqrt{\frac{m\omega}{\pi\hbar}}e^{-m\omega x^2 / \hbar} dx\\
  &= 0.16\\
\end{align*}

Thus we see that there is a small chance of finding the particle in the classically
forbidden region, which is expected for the ground state of the QHO. To do the
calculation we exploit the fact that even functions can be integrated on only one
side of the y-axis and doubled to get the total integral answer.\\

@@

%% \begin{align*}
%%   c_0 &= \int_{-\infty}^{\infty} \phi_0^*\psi dx\\
%%   &= A\int_{-\infty}^{\infty} \left(\frac{m\omega}{\pi\hbar}\right)^{1/2}e^{-m\omega x^2/\hbar}
%%   \left(1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
%%   2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/(2\hbar)}dx\\
%%   &= 2A\\
%%   c_1 &= A\int_{-\infty}^{\infty} \frac{2\sqrt{\frac{m\omega}{\hbar}x}}{\sqrt{2}}
%%     \left(\frac{m\omega}{\pi\hbar}\right)^{1/2} \left(1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
%%     2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/\hbar}dx\\
%%  &= -\frac{3}{\sqrt{2}}A\\
%%  c_2 &= A\int_{-\infty}^{\infty} \frac{\frac{4m\omega}{\hbar}x^2 - 2}{\sqrt{8}}
%%     \left(\frac{m\omega}{\pi\hbar}\right)^{1/2} \left(1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
%%     2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/\hbar}dx\\
%%  &= \sqrt{2}A\\
%% \end{align*}

%% From this we can see that $\psi(x,0) = \sqrt{
%% \left(\frac{2\sqrt{2}}{\sqrt{21}\ket{0}
%%+ \frac{-3}{\sqrt{21}}\ket{1} + \frac{sqrt{2}}{\sqrt{21}}\ket{2}\right)$. It is then easy to compute
%% the expectation value: \\

%% \begin{align*}
%%   <E> &= \sum \hbar\omega\left(n+\frac{1}{2}\right)|c_n|^2\\
%%   &= \frac{2 + \frac{27}{4} + 5}{\frac{21}{2}} = 1.3\hbar\omega\\
%% \end{align*}
\section{9.13: Half-QHO well}
(Inspired by http://www.physicspages.com/2012/08/18/half-harmonic-oscillator/)

This system is identical to the 1D QHO, except all positive x values have infinite
potential. So the system is a mixture of the infinite well and SHO potentials. We
thus expect the wave function to have characteristics of both systems. For positive
x values we expect the wave function to be zero, as it is in the infinite well.
Because the negative x-values have a potential identical to QHO, we expect the wave
function to be identical as well. So our total wave function is a Hermite/Gaussian
for negative x values and zero for positive x values.\\

However, we're not done. We notice that the x=0 boundary condition is not obeyed
for all Hermite/Gaussian wave functions. We require the wave function to be zero
at x=0, which is only true for odd QHO wave functions. Thus our energy spectrum
is identical to the QHO spectru, but only for odd values of n:

\begin{align*}
  E_n &= \left(n + \frac{1}{2}\right)\hbar\omega\hspace{2cm}\mbox{n=1,3,5,...}\\
\end{align*}

For this problem we just applied the basic logic that the value of a wave function
is only based on its current location, not the overall shape of the potential. The
wave function for a system that is identical to the QHO will have a wave function
identical to QHO in the similar regions. This is a powerful result.\\

\section{5: 2D QHO}
(Derivation inspired by
$\mbox{nyu.edu/classes/tuckerman/adv.chem/lectures/lecture\_18/node6.html)}$\\
To take the QHO to two dimensions, we use the following Hamiltonian:

@@

We notice that none of the cartesian terms in this Hamiltonian are coupled, so
our energy eigenvalue equation then takes the form:

\begin{align*}
  H_{xy}X(x)Y(y) &= (E_x + E_y)X(x)Y(y)\\
\end{align*}

@@

\begin{align*}
  \psi_0(x,y) &= \left(\frac{m\omega}{\pi\hbar}\right)^{1/4}
  e^{-m\omega\left(x^2+y^2\right)/2\hbar}\\
\end{align*}

The ease with which we go from the 1D to 2D harmonic oscillators comes from the fact
that the x and y coordinates of the oscillator are not coupled. The x and y cross
sections of the wave can take on any allowed 1D state. It is not intuitive to me
why this method exactly works. Why are the energies of the x and y components
just summed together to get total energy, as opposed to being multiplied or
something? The degeneracy of the system is interesting. Because a sum is used,
for large energies there are a huge number of ways to configure the system.\\
